\chapter{Known Issues \& Technical Debt}
\label{ch:known-issues}

This chapter used to list 21 items — every \textbf{KNOWN ISSUE} flagged earlier in this manual, collected in one place so a maintainer could find "what's currently wrong with this codebase" without reading 21 chapters first. All 21 have since been fixed (correctness bugs, security gaps, and most of the design/consistency debt), each in the chapter that originally flagged it, per this chapter's own house rule below. A pre-existing lint-debt problem discovered along the way (Chapter~\ref{ch:workflow}) has since been fixed too — and so has the 37-warning \code{react-hooks/exhaustive-deps}/\code{@next/next/no-img-element} debt that this chapter went on to flag once the error-level debt was clear, and the stock-Tailwind-to-\code{brand-*} token migration (Chapter~\ref{ch:styling}) this chapter flagged next. What's left is a short, honest list of what's still actually open.

\section{Still open}

\begin{enumerate}
  \item \textbf{\pathname{about-modal.tsx} still has a placeholder bio for one founder.} Owais Saleem's bio is real; Malik Umair's is still "Biography coming soon." — waiting on him to provide the text, not a code task. Chapter~\ref{ch:components}.
  \item \textbf{\code{alert()} is still the normal way most pages report a failed action.} The two instances originally flagged (\pathname{rating-modal.tsx}, \pathname{clubs/[id]}'s delete confirm) are fixed, but \pathname{browse/page.tsx}, \pathname{requests/page.tsx}, \pathname{my-books/page.tsx}, \pathname{messages/[id]/page.tsx}, and \pathname{clubs/create/page.tsx} all still use raw \code{alert()} for error messages. This is a real, larger consistency project — replacing all of them with inline styled messages — not something this pass took on. Chapter~\ref{ch:components}.
  \item \textbf{Automated test coverage is minimal.} Vitest is configured and covers a handful of pure functions (Chapter~\ref{ch:workflow}), but there is no component, integration, or end-to-end coverage. Growing this is ongoing, incremental work, not a single fixable bug.
\end{enumerate}

\begin{olknote}
If you fix one of the items above, delete its entry from this chapter in the same pull request — an item that's been fixed but still listed here is its own small instance of documentation drift, which is exactly the failure mode this chapter exists to avoid repeating.
\end{olknote}
