\chapter{Wishlists, Matching, Clubs \& Events}
\label{ch:wishlists-clubs}

\section{Wishlists: asking for a book that doesn't exist yet}

A wishlist entry (\pathname{wishlist/page.tsx}, Chapter~\ref{ch:pages}) is just a title, author, and genre a user wants but couldn't find on Browse. On its own it does nothing — it becomes useful only through the matching flow triggered from the \emph{other} side of the transaction.

\subsection*{The matching flow}

\begin{enumerate}
  \item User A adds \emph{"The Old Man and the Sea"} to their wishlist. Nothing happens yet — \code{matched\_book\_id} stays \code{NULL}.
  \item User B, independently, lists a copy of \emph{"The Old Man and The Sea"} (different capitalisation, doesn't matter) on \pathname{my-books}.
  \item \pathname{my-books/page.tsx} calls \code{match\_wishlists\_for\_book(title, owner\_id)} (Chapter~\ref{ch:functions}) right after the insert.
  \item The \code{pg\_trgm} similarity search finds User A's wishlist row (similarity comfortably above the 0.35 threshold despite the capitalisation difference), and the frontend updates that row's \code{matched\_book\_id} and fires a notification to User A.
\end{enumerate}

\begin{olknote}
Re-read Chapter~\ref{ch:functions}'s explanation of \emph{why} this needs to be a \code{SECURITY DEFINER} RPC rather than a plain query: the \code{wishlists} RLS policy (Chapter~\ref{ch:rls}) restricts a normal \code{SELECT} to \code{user\_id = auth.uid()}. User B's own session, running the match search, could never see User A's wishlist row through an ordinary query — the whole feature would be silently non-functional without the RPC's elevated privilege, which is exactly the bug this migration fixed.
\end{olknote}

The unique index on \code{(user\_id, lower(btrim(title)))} (Chapter~\ref{ch:schema}) exists to stop a second, independent problem: without it, adding the same title twice (perhaps after a page refresh double-submits a form) would previously have produced two wishlist rows, and a later match would notify the user twice for the same book.

\section{Clubs: geography plus a trust gate}

Clubs are the only feature in the schema besides books themselves that carries its own latitude/longitude (Chapter~\ref{ch:schema}) and its own nearby-search RPC, \code{get\_clubs\_nearby} (Chapter~\ref{ch:functions}) — the same haversine pattern, applied to a different table, ordered by distance then by \code{member\_count}.

\subsection*{Creating a club is gated on trust, in the UI and now the database}

\pathname{clubs/create/page.tsx} (Chapter~\ref{ch:pages}) requires 5+ completed exchanges and zero reports against the creator, computed with two client-side count queries before the form will even submit. The \code{clubs} table's INSERT policy still allows any authenticated user (Chapter~\ref{ch:rls}), but a \code{BEFORE INSERT} trigger, \code{check\_club\_creation\_eligibility} (Chapter~\ref{ch:functions}), now re-derives the same eligibility rule at the database layer, so the gate holds even for a request that bypasses the UI.

\begin{olkhowto}
Raising the bar to, say, 10 exchanges means editing the number in \emph{two} places, both of which currently hard-code \code{5}:
\begin{enumerate}
  \item \pathname{src/app/(dashboard)/clubs/create/page.tsx}, line 81 — \code{setEligible(completedCount >= 5 \&\& !userHasReports)} — plus the \code{5} inside the "5+ completed exchanges" label around line 132, or the badge will contradict the actual gate.
  \item \code{check\_club\_creation\_eligibility()} in \pathname{supabase/schema.sql} — the line \code{IF completed\_count < 5 OR report\_count > 0 THEN}. Never hand-edit \pathname{schema.sql} directly: run \code{npx supabase migration new raise\_club\_threshold}, put the \code{CREATE OR REPLACE FUNCTION} with \code{< 10} in the new migration file, then apply it (Chapter~\ref{ch:local-env}).
\end{enumerate}
Ship only \#1 and the UI merely \emph{feels} stricter — a direct API call still gets through at the old threshold, since the trigger is what actually enforces the rule.
\end{olkhowto}

\subsection*{Membership and posts}

Joining is a simple \code{club\_members} insert, kept honest by \code{trg\_club\_member\_count} (Chapter~\ref{ch:functions}) rather than client-side arithmetic. \code{removeClubMember} in admin actions (Chapter~\ref{ch:admin}) used to redundantly re-count on top of the trigger, introducing an off-by-one on every admin-initiated removal; that manual re-count has been removed.

Posting to a club is creator-only (Chapter~\ref{ch:rls}); a new post fans out one notification per member via a single batched \code{.insert(notifications[])} call (Chapter~\ref{ch:pages}) — the same N+1-avoidance discipline seen in the messages list (Chapter~\ref{ch:pages}) and the admin broadcast action (Chapter~\ref{ch:admin}).

\subsection*{Deactivation is soft, always}

Neither a club's creator nor an admin can hard-delete a club through the app layer — \code{active = false} is the only operation exposed (\pathname{clubs/[id]/page.tsx}'s "delete," and the admin \code{deactivateClub}/\code{reactivateClub} actions, Chapter~\ref{ch:admin}). Only a \code{super\_admin}, and only directly via SQL, could actually remove a \code{clubs} row (Chapter~\ref{ch:rls}'s DELETE policy) — there is no UI path to a hard delete at all, by design.

\begin{olktask}
Add a wishlist entry for a title that doesn't exist yet, then list a book with a \emph{deliberately misspelled} version of that title from a second test account (swap two letters, or use a different capitalisation) and confirm the match still fires. Then try a title that only shares one common word (e.g. "The Old Man" vs. "Old Man River") and confirm it does \emph{not} match — this is the fastest way to build a feel for where the 0.35 similarity threshold actually sits in practice, beyond just reading the number.
\end{olktask}

\section{Events: club-scoped, but publicly discoverable}

A club can schedule events (\code{club\_events}, plus one \code{event\_rsvps} row per attendee — Chapter~\ref{ch:schema}): a title, description, start/end time, either an in-person location or an online meeting link, an optional attendee cap, and a per-event \code{visibility} of \code{public} or \code{members\_only}.

\begin{olknote}
\textbf{\code{visibility} gates the RSVP action, not discovery.} Every event stays listed and viewable by anyone — on the club's own page, on the cross-club \pathname{/events} directory, and in the landing page's "Upcoming Events" preview — regardless of the viewer's club membership, mirroring how clubs themselves are already publicly browsable (Chapter~\ref{ch:rls}). A \code{members\_only} event just shows "Join the club to RSVP" instead of an RSVP button for a non-member. Two different features, deliberately not conflated: hiding an event outright would need a second, separate flag.
\end{olknote}

Creating an event is gated exactly like posting an announcement (Chapter~\ref{ch:rls}) — only the club's creator — rather than the club-creation eligibility rule earlier in this chapter; a club's creator has already cleared that bar once, so there is no second trigger to re-derive it for events. \code{trg\_event\_attendee\_count} keeps \code{club\_events.attendee\_count} in sync with real \code{event\_rsvps} rows, the same atomic-trigger shape as \code{trg\_club\_member\_count} above (Chapter~\ref{ch:functions}).

\begin{olkgotcha}
The RSVP \code{INSERT} policy's \code{visibility} check was added alongside a second, independent condition — \code{capacity IS NULL OR attendee\_count < capacity} — but that second condition didn't ship on day one. The "Event full" state was originally enforced only by disabling the RSVP button once \code{attendee\_count >= capacity} in \pathname{events/[id]/page.tsx}; a direct \code{POST /rest/v1/event\_rsvps} call could still RSVP past a full event, confirmed by testing exactly that against a capacity-1 event before the policy was corrected. Same lesson as the club-creation gate earlier in this chapter, one feature later: a disabled button is a UX nicety, not a security boundary — only the policy is.
\end{olkgotcha}

Event notifications reuse the club-announcement broadcast shape: creating an event fans out an \code{event\_created} notification to every current member (Chapter~\ref{ch:server-actions}), routed through the same \code{can\_notify} guard with a new branch verifying the caller is the event's own creator and the target is a real club member. There is deliberately no per-RSVP notification to the organiser — at any real scale that would be pure noise, so the attendee list on the event and admin pages covers that need instead.

The cross-club directory (\pathname{/events}) uses \code{get\_events\_nearby}, the same haversine-and-\code{NULLS LAST} pattern as \code{get\_clubs\_nearby} (Chapter~\ref{ch:functions}), falling back to a plain \code{created\_at}-ordered query when the viewer has no location set. A zero-dependency Route Handler, \pathname{api/events/[id]/ics/route.ts}, generates an RFC5545 calendar file on demand for "Add to calendar" — no external library, just a template string with the handful of characters (\code{,}, \code{;}, \code{\textbackslash}) that need escaping inside an ICS field.

\begin{olktask}
Create a club event with \code{capacity} set to 1, RSVP to it yourself, then — from a second test account that is \emph{not} a member of that club — try to RSVP through the UI (should be blocked by the disabled button) and then directly via a \code{POST} to \pathname{/rest/v1/event\_rsvps} with that account's own access token (should come back \code{403}). This is the same "is it a real gate or just a hidden button" habit Chapter~\ref{ch:rls}'s task asks you to build for club creation, applied to the newest feature in the schema.
\end{olktask}
