\chapter{Known Issues \& Technical Debt}
\label{ch:known-issues}

This chapter used to list 21 items — every \textbf{KNOWN ISSUE} flagged earlier in this manual, collected in one place so a maintainer could find "what's currently wrong with this codebase" without reading 21 chapters first. All 21 have since been fixed (correctness bugs, security gaps, and most of the design/consistency debt), each in the chapter that originally flagged it, per this chapter's own house rule below. A pre-existing lint-debt problem discovered along the way (Chapter~\ref{ch:workflow}) has since been fixed too — and so has the 37-warning \code{react-hooks/exhaustive-deps}/\code{@next/next/no-img-element} debt that this chapter went on to flag once the error-level debt was clear, and the stock-Tailwind-to-\code{brand-*} token migration (Chapter~\ref{ch:styling}) this chapter flagged next. The remaining \code{alert()} consistency item below has since been closed out too, by a full security \& code-quality audit (Chapter~\ref{ch:playbook}) that also fixed several RLS gaps (Chapter~\ref{ch:rls}) and a notification-spoofing hole (Chapter~\ref{ch:server-actions}) that this chapter had never flagged in the first place — this manual only documents debt once someone finds it, not debt that exists but hasn't been noticed yet. What's left is a short, honest list of what's still actually open.

\section{Still open}

\begin{enumerate}
  \item \textbf{\pathname{about-modal.tsx} still has a placeholder bio for one founder.} Owais Saleem's bio is real; Malik Umair's is still "Biography coming soon." — waiting on him to provide the text, not a code task. Chapter~\ref{ch:components}.
  \item \textbf{Automated test coverage is minimal.} Vitest is configured and covers a handful of pure functions (Chapter~\ref{ch:workflow}), but there is no component, integration, or end-to-end coverage. Growing this is ongoing, incremental work, not a single fixable bug.
  \item \textbf{No error monitoring in production.} No Sentry, no Vercel Analytics/Speed Insights, nothing — a runtime error today is invisible unless a user reports it. Chapter~\ref{ch:welcome}'s scaling roadmap (Phase 1) lists this as a prerequisite for real growth, not yet done.
  \item \textbf{No general HTTP-level rate limiting.} The only rate limits that exist are two database triggers scoped narrowly to \code{messages} and \code{book\_requests} (Chapter~\ref{ch:functions}). Login, signup, and plain \code{/browse} traffic are unprotected. Chapter~\ref{ch:welcome}, Phase 1.
  \item \textbf{\pathname{browse/page.tsx}'s \code{.limit(200)} and \code{get\_books\_nearby}'s \code{p\_limit} are stopgaps, not real pagination.} Once the catalog exceeds a couple hundred available books, results past that cutoff silently stop appearing — no error, no "next page." Chapter~\ref{ch:welcome}, Phase 1.
  \item \textbf{Admin list pages other than Books have no pagination.} Users, Requests, Reports, Clubs, Events, and Club Requests all fetch their full result set. Lower priority than the item above since admin traffic is inherently small, but will degrade as those tables grow over months of real usage. Chapter~\ref{ch:pages}.
\end{enumerate}

\begin{olknote}
If you fix one of the items above, delete its entry from this chapter in the same pull request — an item that's been fixed but still listed here is its own small instance of documentation drift, which is exactly the failure mode this chapter exists to avoid repeating.
\end{olknote}
